New lower bounds for Schur and weak Schur numbers

### 2021-09-20 article/weakschur.tex
valeurs bis
contexte + titre 3.5

Inequalities (\ref{WS(n+1)}) and (\ref{WS(n+2)}) were found by Rowley, they are detailed in \cite{RowleyWS}. 
Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in \hyperref[ConstructionWS]{Subsection 3.3} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that bettert WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

@@

\label{LowerBoundsWS}
\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-4}
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-2}
		\cellcolor{yellow}{4-3}
	\Body
		n & 8 & 9 & 10 & 11 \\
		4 \, S(n-1) + 2 & 6\,722 & 21\,146 & 71\,214 & 243\,794 \\
		13 \, S(n-2) + 8 & 6\,976 & 21\,848 & 68\,726 & 231\,447 \\
		42 \, S(n-3) + 24 & 6\,744 & 22\,536 & 70\,584 & 222\,036 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
	\Body
		n & 12 & 13 & 14 & 15 \\
		4 \, S(n-1) + 2 & 815\,314 & 2\,554\,194 & 8\,045\,162 & 27\,061\,154 \\
		13 \, S(n-2) + 8 & 792\,332 & 2\,649\,772 & 8\,301\,132 & 26\,146\,778 \\
		42 \, S(n-3) + 24 & 747\,750 & 2\,559\,840 & 8\,560\,800 &  25\,886\,224 \\
	\end{NiceArray}
\end{array}
\]
\caption{New lower bounds for \( n \in [\![8,15]\!] \)}
\end{table}

@@

With \( S(9) \geqslant 17\,803 \), we found a new lower bound for \(\WS(10)\) using (\ref{WS(n+1)}).
Moreover, inequality (\ref{WS(n+3)}) gives new lower bounds for \(\WS(9)\) and \(WS(14)\).

\subsection{\(\WS(6) \geqslant 646\)}
\qquad When applying the inequality \(\WS(n+1) \geqslant 4 \, S(n) + 2\), one may realize that it is possible to build a weakly 
sum-free partition of length \(4 \, S(n) + 3\) for small values of \(n\) (\(n \leqslant 4\)) by using the contruction of 
\hyperref[Corollary3.19]{Corollary 3.19} for the integers 1, 2, \(4i\) and \(4i+1\) for \(i \in [\![1, S(n)]\!]\) but not 
constraining the other integers. We did the same for the Schur partitions corresponding to Schur number five. More precisely, 
we imposed these constraints only for \(i \leqslant 50\) in order to have more degrees of freedom. The number 50 was chosen 
arbitrarly so that all of the Schur number 5-partitions could be tested in a few hours.

>>>>

Inequalities (\ref{WS(n+1)}) and (\ref{WS(n+2)}) were found by Rowley, they are detailed in \cite{RowleyWS}. 
Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in \hyperref[ConstructionWS]{Subsection 3.3} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

@@

\label{LowerBoundsWS}
\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-5}
		\cellcolor{yellow}{3-2}
		\cellcolor{yellow}{4-3}
		\cellcolor{yellow}{4-4}
	\Body
		n & 8 & 9 & 10 & 11 \\
		4 \, S(n-1) + 2 & 6\,786 & 21\,146 & 71\,214 & 243\,794 \\
		13 \, S(n-2) + 8 & 6\,976 & 22\,056 & 68\,726 & 231\,447 \\
		42 \, S(n-3) + 24 & 6\,744 & 22\,536 & 71\,256 & 222\,036 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 12 & 13 & 14 & 15 \\
		4 \, S(n-1) + 2 & 815\,314 & 2\,578\,514 & 8\,045\,162 & 27\,061\,154 \\
		13 \, S(n-2) + 8 & 792\,332 & 2\,649\,772 & 8\,380\,172 & 26\,146\,778 \\
		42 \, S(n-3) + 24 & 747\,750 & 2\,559\,840 & 8\,560\,800 &  27\,074\,400 \\
	\end{NiceArray}
\end{array}
\]
\caption{New lower bounds for \( n \in [\![8,15]\!] \)}
\end{table}

@@

With \( S(9) \geqslant 17\,803 \), we found a new lower bound for \(\WS(10)\) using (\ref{WS(n+1)}).
Moreover, inequality (\ref{WS(n+3)}) gives new lower bounds for \(\WS(9)\), \(\WS(10)\), \(\WS(14)\) and \(\WS(15)\).

\subsection{Another approach: \(\WS(6) \geqslant 646\)}
\qquad The weak Schur partitions obtained with WS-templates have an extremely regular structure. Therefore, one may 
expect to find larger partitions by allowing more freedom in the structure of the partition while still preserving a structure 
somewhat close to a template-based partition thus reducing the search space to a manageable size.

When applying the inequality \(\WS(n+1) \geqslant 4 \, S(n) + 2\), one may realize that it is possible to build a weakly 
sum-free partition of length \(4 \, S(n) + 3\) for small values of \(n\) (\(n \leqslant 4\)) by using the contruction of 
\hyperref[Corollary3.19]{Corollary 3.19} for the integers 1, 2, \(4i\) and \(4i+1\) for \(i \in [\![1, S(n)]\!]\) but not 
constraining the other integers. We did the same for the Schur partitions corresponding to Schur number five. More precisely, 
we imposed these constraints only for \(i \leqslant 50\) in order to have more degrees of freedom. The number 50 was chosen 
arbitrarly so that all of the Schur number 5-partitions could be tested in a few hours.
